---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


## Loading and preprocessing the data

```{r}
data = read.csv("activity.csv")
head(data)
class(data)
data.clean = data[complete.cases(data), ]
```

## What is mean total number of steps taken per day?

a histogram of the total number of steps taken each day

```{r, echo=TRUE}
library("ggplot2")
g = ggplot(data.clean, aes(steps))
g+geom_histogram(binwidth = 30)
```

mean and median total number of steps taken per day

```{r}
Data_day = aggregate(steps ~ date, data=data, FUN = "sum")
print(mean(Data_day$steps))
print(median(Data_day$steps))
```


## What is the average daily activity pattern?

calculating the average number of steps taken, averaged across all days

```{r}
data.average <- aggregate(steps ~ interval, data=data.clean, FUN = "mean")
head(data.average)
```

ploting the data
```{r, echo=TRUE}
p <- ggplot(data.average, aes(x= interval, y=steps))
p+geom_line(col="red")
```

interval, on average across all the days in the dataset,contains the maximum number of steps

```{r, echo=TRUE}
data.average[data.average$steps==max(data.average$steps),]$interval
```


## Imputing missing values
number of missing values
total
```{r}
sum(!complete.cases(data))
```
in steps column
```{r}
sum(is.na(data$steps))
```
in date column
```{r}
sum(is.na(data$date))
```
in interval column 
```{r}
sum(is.na(data$interval))
```
filling the missing values
```{r}
data.final = data.frame(steps=integer(0), date=character(0), interval=integer(0))

for (i in unique(data$interval)) {
    data.NA <- subset(data, interval==i)
    mean.NA = mean(data.NA$steps[!is.na(data.NA$steps)])
    data.NA[is.na(data.NA)]<-mean.NA
    data.final <- rbind(data.final, data.NA)
    }
```

a histogram of the total number of steps taken each day
```{r}
library(ggplot2)
h<-ggplot(data.final, aes(steps))
h+geom_histogram()
```
calculating mean and median
```{r}
Data_day.final = aggregate(steps ~ date, data=data.final, FUN = "sum")
print(mean(Data_day.final$steps))
print(median(Data_day.final$steps))
```


## Are there differences in activity patterns between weekdays and weekends?

converting date column to date format
```{r}
data$date <- as.Date(data$date, format="%Y-%m-%d")
class(data.final$date)

library(chron)

data.weekend <- data.frame(data, weekend = is.weekend(data$date))

Avr.weekday <- aggregate(steps ~ interval + weekend, data=data.weekend, FUN = "mean")
```
plotting weekday data
```{r}
library(ggplot2)
k<-ggplot(Avr.weekday, aes(x=interval, y=steps))
k+geom_line(col="red")+facet_grid(weekend~.)+
  labs(title = "Number of steps between weekdays and weekends")
```






